Hockey Enforcers

### 2023-06-29 EDA.Rmd
commit.

#dead <- do.call(rbind, deaths)
#write.csv(dead, file = "/Users/gregorymatthews/Dropbox/hockeyEnforcers/deaths.csv")
dead <- read.csv("/Users/gregorymatthews/Dropbox/hockeyEnforcers/deaths.csv")
dead <- dead %>% filter(From >= 1967 & Name != "Name")

>>>>

#dead <- do.call(rbind, deaths)
#write.csv(dead, file = "/Users/gregorymatthews/Dropbox/hockeyEnforcers/deaths.csv")
dead <- read.csv("/Users/gregorymatthews/Dropbox/hockeyEnforcersGit/deaths.csv")
dead <- dead %>% filter(From >= 1967 & Name != "Name")

### 2023-06-29 HocketEnforcersReport.Rmd
commit.

```{r}
library(survival)
#Defining the survival variable
dat$Surv <- Surv(dat$age,dat$dead)
```

```{r}
ggplot(aes(x = age,col = EF), data = dat) + geom_density()
ggplot(aes(x = Wt,col = EF), data = dat) + geom_density()
ggplot(aes(x = Wt,y = age,col = EF), data = dat) + geom_point() + geom_smooth()
```

```{r}
ggplot(aes(x = Total.Fights,y = age), data = dat) + geom_point() + geom_smooth()
ggplot(aes(x = Total.Fights,y = age, col = dead), data = dat) + geom_point() + geom_smooth()
ggplot(aes(y = Total.Fights,x = Wt), data = dat) + geom_point() + geom_smooth()
```

@@

# Cox model with Total.fights as a covariate.  
```{r}
a <- coxph(Surv ~ Total.Fights + GP.x + Ht + Wt + DOB.Year + Pos , data = dat)
summary(a)
```
# Cox model with EF as a covariate.  
```{r}
a <- coxph(Surv ~ EF + GP.x + Ht + Wt + DOB.Year + Pos , data = dat)
summary(a)
```

@@

```{r}
coxph(Surv ~ EF + GP.x + Ht + Wt + DOB.Year + Pos, data = dataMatched)
```
# Conditioning on dead people 
All the significance disappears.  
```{r}
dataMatched %>% filter(dead) %>% group_by(EF) %>% summarise(mnage = mean(as.numeric(age)), mnwt = mean(Wt))

@@

@@

b <- lm(age ~ EF + Wt, data = dataMatched[dataMatched$dead,])
summary(b)
```
# Various stuff
```{r}
dataMatched$age <- as.numeric(dataMatched$age)
b <- lm(age ~ Wt, data = dataMatched[dataMatched$dead,])
summary(b)

#Relationship between weight and age for those who are dead. 
ggplot(aes(x = Wt, y = age), data = dataMatched[dataMatched$dead,]) + geom_point() + geom_smooth()

# 
# library(survival)
# dataMatched$Surv <- Surv(dataMatched$age,dataMatched$dead)
# 
# #How many games should a playewr have played to be included????
# coxph(Surv ~ EF, data = dataMatched)
# coxph(Surv ~ EF + GP.x + Ht + Wt + DOB.Year + Pos, data = dataMatched)
# 
# sub <- dataMatched[dataMatched$dead,]
# ggplot(aes(y = age, x = factor(EF)), data = sub) + geom_boxplot()
# 
# sub %>% group_by(EF) %>% summarize(mn = mean(age))
# t.test(sub$age ~ sub$EF)
# sub$age <- as.numeric(sub$age)
# a <- lm(age ~ EF + Ht + Wt + GP.x + Pos, data = sub)
# summary(a)
# 
# 
# ggplot(aes(y = Wt, x = factor(EF)), data = sub) + geom_boxplot()
# ggplot(aes(y = Wt, x = factor(EF)), data = dataMatched) + geom_boxplot()
# 
# 
# 
# 
# 
# coxph(Surv ~ Total.Fights  + GP.x + Ht + Wt + DOB.Year + Pos , data = dat[dat$GP.x >= 10,])
# 
# coxph(Surv ~ I(log(Total.Fights+1)) + GP.x, data = dat[dat$GP.x >= 10,])
# 
# 
# summary(dat$Total.Fights)
# 
# coxph(Surv ~ Total.Fights  + GP + Ht + Wt + DOB.Year + Pos , data = dat[dat$Total.Fights <= 100,])
# 
# km_fit <- survfit(Surv(dat$age,dat$dead) ~ 1)
# plot(km_fit)
# 
# library(ggplot2)
# ggplot(aes(x = Total.Fights, y = age, col = dead), data = dat[dat$Total.Fights <= 100,]) + geom_point() + geom_smooth()
# dat %>% group_by(DOB.Year) %>% summarise(fperg = sum(Total.Fights)/sum(GP)) %>% ggplot(aes(x = DOB.Year, y = fperg)) + geom_point() + geom_smooth()
# dat %>% group_by(X1st.Season.Corrected) %>% summarise(fperg = sum(Total.Fights)/sum(GP)) %>% ggplot(aes(x = X1st.Season.Corrected, y = fperg)) + geom_point() + geom_smooth()
# 
# ggplot(aes(x = Total.Fights, y = age, col = dead), data = dat[dat$Total.Fights <= 100,]) + geom_point() + geom_smooth()
# 
#

@@

# Simulating from a coxph model with a Weibull baseline
# Everything below here is still nonsense.  
This is far from complete....
```{r}

#https://mathildesautreuil.github.io/survMS/articles/how-to-simulate-survival-models.html
#mean = 1/lambda*gamma(1+a)
lambda <- 4000000
a <- -3.8
1/(lambda^(1/a))*gamma(1+1/a)
n <- 1000
x <- runif(n/2,0,10)
dat.sim <- data.frame(1,x = rep(x,2), EF = rep(c(0,1),each = n/2))
dat.sim$Wt <- rnorm(n,200,15) 
dat.sim$Wt[dat.sim$EF == 1] <- rnorm(n/2,215,15) 
beta <- c(0,0,0,0.05)
U <- runif(n)
times <- 1/(lambda^(1/a))*((-log(1-U)/(exp(as.matrix(dat.sim)%*%beta))))^(1/a)
mean(times)
hist(times)
summary(times)

dat.sim$times <- times

dat.sim %>% group_by(EF) %>% summarise(mean = mean(times), n = n())

dat.sim %>% filter(times <= 951) %>% group_by(EF) %>% summarise(mean = mean(times), n = n())

```{r}
set.seed(1234)
weightE <- rnorm(100, 215, 15)
weightC <- rnorm(100, 200, 15)

x <- data.frame(Wt = c(weightE, weightC), EF = rep(c(1,0),each = 100))
eps <- rnorm(100,0,3)
x$age <- 60 + 0.5*(x$Wt - 200) + eps
x$age <- 60 - 0.5*(x$Wt - 200) + eps
summary(x$age)
x$dead <- rbinom(200,1,0.2)

library(survival)
x$age[x$dead == 0] <- x$age[x$dead == 0] - runif(sum(x$dead == 0),0,30)
coxph(Surv(age,dead) ~ EF, data = x)
table(x$dead,x$EF)
prop.test(c(29,18),c(100,100))

x %>% filter(dead == 1) %>% group_by(EF) %>% summarise(mn = mean(age))
t.test(x$age[x$dead == 1] ~ x$EF[x$dead == 1])

>>>>

Creating the Surv object
```{r}
library(survival)
#Defining the survival variable
dat$Surv <- Surv(dat$age,dat$dead)
```
Some plots.
```{r}
ggplot(aes(x = age,col = EF), data = dat) + geom_density()
ggplot(aes(x = Wt,col = EF), data = dat) + geom_density()
ggplot(aes(x = Wt,y = age,col = EF), data = dat) + geom_point() + geom_smooth()
```
Some more plots.
```{r}
ggplot(aes(x = Total.Fights,y = age), data = dat) + geom_point() + geom_smooth()
ggplot(aes(x = Total.Fights,y = age, col = dead), data = dat) + geom_point() + geom_smooth()
ggplot(aes(y = Total.Fights,x = Wt), data = dat) + geom_point() + geom_smooth()
```

@@

# Cox model with Total.fights as a covariate.  
```{r}
a <- coxph(Surv ~ Total.Fights + GP.x + Ht + Wt + DOB.Year + Pos , data = dat)
summary(a)
```
# Cox model with EF as a covariate.  
No significant effect for enforcers in terms of survival times.  
Though the sign of the effect is what you would expect to see if there was a risk to being an enforcer.

```{r}
a <- coxph(Surv ~ EF + GP.x + Ht + Wt + DOB.Year + Pos , data = dat)
summary(a)
```

@@

```{r}
coxph(Surv ~ EF + GP.x + Ht + Wt + DOB.Year + Pos, data = dataMatched)
```
# Conditioning on dead people

```{r}
dataMatched %>% filter(dead) %>% group_by(EF) %>% summarise(mnage = mean(as.numeric(age)), mnwt = mean(Wt), mnDOB = mean(DOB.Year), mngp = mean(GP.x))

@@

t.test(as.numeric(dataMatched$GP.x[dataMatched$dead]) ~ dataMatched$EF[dataMatched$dead])

t.test(as.numeric(dataMatched$DOB.Year[dataMatched$dead]) ~ dataMatched$EF[dataMatched$dead])

@@

b <- lm(age ~ EF + Wt, data = dataMatched[dataMatched$dead,])
summary(b)

#What does a survival analysis look like conditional on death?
test <- dataMatched %>% filter(dead)
coxph(Surv ~ EF + GP.x + Ht + Wt + DOB.Year + Pos, data = test)

#Pick something else to stratify on and you get absurd answer
#As an example, filter out only those players who have played more than 700 games.
#The age in the enforcers group is significantly larger than the age in the non-enforcers group.  
pd <- 700
dataMatched %>% filter(GP.x > pd) %>% group_by(EF) %>% summarise(n = n(),mnage = mean(as.numeric(age)), mnwt = mean(Wt), mnDOB = mean(DOB.Year), mngp = mean(GP.x))

t.test(as.numeric(dataMatched$age[dataMatched$GP.x > pd]) ~ dataMatched$EF[dataMatched$GP.x > pd])

@@
